---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo = TRUE}
library(ggplot2)
activity <- read.csv("activity.csv")
```

## Create Histogram
```{r, echo = TRUE}
steps_per_day<- aggregate(steps ~ date, activity, FUN=sum)
Histogram1 <- hist(steps_per_day$steps, col = "green", main = "Steps Per Day", xlab = "Steps")
Histogram1
```

## What is mean and median total number of steps taken per day?
```{r, echo = TRUE}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```

## What is the average daily activity pattern?
```{r, echo = TRUE}
daily_activity <- aggregate(steps ~ interval, activity, FUN = mean, na.rm = TRUE)
names(daily_activity) <- c("interval", "avgsteps")
Plot1 <- ggplot(daily_activity, aes(interval, avgsteps)) + geom_line() + ggtitle("Average Daily Activity Pattern") + xlab("5-min Intervals") + ylab("Average Steps per Day")
Plot1
```

## Which 5-min interval contains the max number of steps?
```{r, echo = TRUE}
daily_activity[which.max(daily_activity$avgsteps), "interval"]
```


## Total number of missing values
```{r, echo=TRUE}
sum(is.na(activity$steps))
```

##Fill in missing values with average steps per day
```{r, echo=TRUE}
ReplacedNA <- activity
ReplacedNA$steps[is.na(ReplacedNA$steps)] <- mean(daily_activity$avgsteps)
```

##Histogram and mean/median of new dataset without missing values
```{r, echo=TRUE}
steps_per_day2 <- aggregate(steps ~ date, ReplacedNA, FUN = sum, na.rm = TRUE)
Histogram2 <- hist(steps_per_day2$steps, col = "green", main = "Steps Per Day", xlab = "Steps")
Histogram2

mean(steps_per_day2$steps)
median(steps_per_day2$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
ReplacedNA$date <- as.Date(strptime(ReplacedNA$date, format="%Y-%m-%d"))
days <- weekdays(ReplacedNA$date)
DayofWeek <-sapply(ReplacedNA$date, function(x) {
    if(weekdays(x) == "Saturday" | weekdays(x) == "Sunday")
    {day <- "Weekend"}
    else {day <- "Weekday"}
    day })

ReplacedNA <- cbind(ReplacedNA, days, DayofWeek)

daytype_activity <- aggregate(steps ~ interval + DayofWeek, ReplacedNA, FUN = mean, na.rm = TRUE)
names(daytype_activity) <- c("interval", "DayofWeek", "avgsteps")
Plot2 <- ggplot(daytype_activity, aes(interval, avgsteps, color = DayofWeek)) + geom_line() + facet_wrap(~DayofWeek) + ggtitle("Daily Average Steps by Day of Week") + xlab("Interval") + ylab("Average Number of Steps")
Plot2
```